---
title: "Analysis_1_SNPS_CS"
author: "William Denault"
date: "2025-06-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 
 Here we have a look  at the 1 SNP CS reported by fSuSiE in the fungen xqtl 
 study

```{r load_data}
library(cowplot)
library(tidyverse)
library(data.table) 
 out <- fread("../data/fsusie_LD.tsv")
 
```



Fsusie was applied to H3K9ac and to DNAm we start by investigating the 1 SNP CS in DNAm.
Then we have a look at H3K9ac.


## Definition Hamming distance

The Hamming distance between two vectors say $x_1=  (a,b,c,d,e)$ and  $x_2=  (a,c,c,d,e)$ 

is the number of bit in which $x_1$ and $x_2$ differs, so in our case above the Hamming distance
between $x_1$ and $x_2$ is 1. In the case  $x_1=  (a,b,c,d,e)$ and  $x_2=  (v,c,c,d,e)$ 
the Hamming distance between $x_1$ and $x_2$ is 2.

## DNAm


First we start by looking at the 1 SNP CS and what is the distance to the Hamming distance of the  closest SNP in the region.
In other what is SNP that is the most similar to the 1-SNP-CS and by "how many individual" do they differ.




```{r}
library(tidyverse)
library(scales)
library(cowplot)
plot_df=out
haQTL_df0 =plot_df%>% filter()%>%filter(str_detect(cs_id,"haQTL") )
 plot(c(table(haQTL_df0$N_diff_samples)  )/length(haQTL_df0$N_diff_samples)*100 ,
     xlab="Hamming distance to nearest SNP",
     ylab ="% of 1-SNP-CS",
     pch=19)

haQTL_df =plot_df%>% filter()%>%filter(str_detect(cs_id,"haQTL"),N_diff_samples< 31)
#percentage of 1 SNP CS in haQTL
table(haQTL_df$N_diff_samples)[1:6]/nrow(haQTL_df0)

ggplot(haQTL_df, aes( x= as.factor(N_diff_samples),
                      y=-log10(haQTL_df$min_P)))+
  geom_boxplot()
```
For haQTL we have about 6 % of 1 SNP CS
with a -log10 pvalue of 

```{r}
quantile(-log10(haQTL_df$min_P[which(haQTL_df$N_diff_samples==1)]))
```

which is relatively high level  of significance


### DNAm

```{r}
mQTL_df0 = plot_df%>% filter()%>%filter(str_detect(cs_id,"mQTL") )
mQTL_df = plot_df%>% filter()%>%filter(str_detect(cs_id,"mQTL"),N_diff_samples< 31)
table(mQTL_df$N_diff_samples)[1:5]/ nrow(mQTL_df0)


ggplot(mQTL_df, aes( x= top_R2))+
  geom_histogram(bins = 100)+
  xlim(c(0.99,1.0002))

ggplot(mQTL_df, aes( x= as.factor(N_diff_samples),
                      y=-log10(mQTL_df$min_P)))+
  geom_boxplot()

```


So for DNAm of we have a much higher number but we also see very strong association
with a lower quantile 

```{r}
quantile(-log10(mQTL_df$min_P[which(mQTL_df$N_diff_samples==1)]))
```

we have a larger percentage for DNAm but the 
```{r}
 
ggplot(tm[ which(tm$people<30),], aes(x=as.factor(people), y= -log10(min_pval)))+
  xlab("Hamming distance")+
    scale_y_continuous(breaks = c(0, 50, 100, 150, 200, 250, 300)) +
  geom_boxplot() + 
  theme_cowplot()
```


 
 
 
 
 
```{r}

n_people_ha= out$ha_long_ratio%>%mutate(people = ceiling (ratio  * 592 +1) )%>%filter() 
tha= merge(n_people_ha  ,out$ha, by= "cs_id")
 
table(n_people_ha$people)[1:10] 
length(n_people_ha$people)
```
 
 
```{r}

 
plot(c(table(n_people_ha$people)  )/length(n_people_ha$people)*100 ,
     xlab="Hamming distance to nearest SNP",
     ylab ="% of 1-SNP-CS",
     pch=19)
```

```{r}
 
 
```