---
title: Illustration of spatial structure in WGBS methylation and H3K27ac ChIP-seq
author: Peter Carbonetto
output: workflowr::wflow_html
---

*Add text here.*

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load the example DNA methylation (WGBS) and H3K9ac (ChIP-seq) signals.

```{r load-data}
load("../data/smash_meth_ha.RData")
```

Load smashr:

```{r load-pkgs}
library(smashr)
```

Set the seed to make the results reproducible:

```{r set-seed}
set.seed(1)
```

WGBS DNA methylation example
----------------------------

This plot shows the beta values at the CpG locations for a single
sample (black dots) and the "smoothed" signal obtained by running
SMASH on these data (blue curve):

```{r methyl-plot, fig.height=2.75, fig.width=7}
y <- methyl_dat$beta
res <- smash.gaus(y,v.est = FALSE)
plot(y,pch = 20,col = "black",cex = 0.65,xlab = "CpG",ylab = "beta value")
lines(res,type = "l",col = "royalblue")
```

Note that the CpG locations are ordered sequencing by base-pair
position along the chromosome (specifically, chromosome 20, near gene
*CASS4*). However, the positions of the CpG probes are not distributed
uniformly; there are some large gaps where there is no probe. The
methylation data are shown in this way to make the spatial structure
more obvious.

ChIP-seq H3K9ac example
-----------------------

This plot shows the read count data from the ChIP-seq assay (black
dots) and the "smoothed" signal obtained by running SMASH on these
data (blue curve):

```{r h3k9ac-plot, fig.height=2.75, fig.width=7}
y <- h3k9ac_dat$y
pos <- h3k9ac_dat$pos
res <- smash.poiss(y,post.var = FALSE)
plot(pos/1e6,y,pch = 20,col = "black",
     xlab = "base-pair position on chromosome 1 (Mb)",
     ylab = "read count")
lines(pos/1e6,res,type = "l",col = "royalblue")
```

This the H3K9ac signal from a region on chromosome 1 near genes *CR1*
and *CR2*.

Note that since these are counts, we used the Poisson-based SMASH
method. Also note that the read counts are actually read counts
averaged across a few samples, so the read counts are not actually
integers.
