---
title: Illustrate whitening property of wavelets
author: Peter Carbonetto
output: workflowr::wflow_html
---

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load the wavethresh package as well as a few other packages used this
demo:

```{r load-pkgs, messages=FALSE, results="hide"}
library(wavethresh)
library(ggplot2)
library(cowplot)
```

Set the seed for reproducibility:

```{r set-seed}
set.seed(1)
```

Simulate a set of signals with ARMA noise:

```{r sim-amira}
n <- 1024
m <- 100
SNR <- 2
v <- DJ.EX()
x <- seq(1,n)/n
ssig <- sd(v$bumps)
sigma <- ssig/SNR
Y <- matrix(0,m,n)
for (i in 1:m) {
  e <- arima.sim(n = n,model = list(ar = 0.99,ma = 1))
  e <- sigma*e/sqrt(var(e))
  y <- v$bumps + e
  Y[i,] <- y
}
```

This is the first noisy signal:

```{r plot-amira, fig.height=2, fig.width=3.5}
y <- Y[1,]
pdat <- data.frame(x = x,y = y)
ggplot(pdat,aes(x = x,y = y)) +
  geom_line() +
  theme_cowplot(font_size = 12)
```

Now let's compute the discrete wavelet transform (DWT) for each
signal:

```{r wd}
D <- matrix(0,m,n-1)
for (i in 1:m) {
  y <- Y[i,]
  D[i,] <- wd(y)$D
}
```

One way to illustrate the decorrelating effect of the wavelet
transform is to compare the autocorrelation for the original signal
and for the wavelet coefficients (WCs). Let's try this on the first
noisy signal:

```{r acf, fig.height=3, fig.width=6}
par(mfrow = c(1,2))
y <- Y[1,]
d <- D[1,]
acf(y)
acf(d)
```

Observe that the autocorrelation is mostly near zero for the WCs
(right).

Another way to illustrate the decorrelating effect of the DWT is to
compute the sample correlations:

```{r cor, fig.height=4, fig.width=3}
cY <- cor(Y)
cD <- cor(D)
p1 <- ggplot(data.frame(x = cY[upper.tri(cY)]),aes(x = x)) +
  geom_histogram(fill = "black",color = "white",bins = 64) +
  xlim(-1,1) +
  labs(x = "correlation",title = "data space") +
  theme_cowplot(font_size = 12)
p2 <- ggplot(data.frame(x = cD[upper.tri(cD)]),aes(x = x)) +
  geom_histogram(fill = "black",color = "white",bins = 64) +
  xlim(-1,1) +
  labs(x = "correlation",title = "wavelet space") +
  theme_cowplot(font_size = 12)
plot_grid(p1,p2,nrow = 2,ncol = 1)
```

Indeed, the strong correlations in the original data space are removed
in the wavelet space.
