---
title: Illustrate whitening property of wavelets
author: Peter Carbonetto
output: workflowr::wflow_html
---

ADD TEXT HERE

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load the wavethresh package as well as a few other packages used this
demo:

```{r load-pkgs, messages=FALSE, results="hide"}
library(wavethresh)
library(ggplot2)
library(cowplot)
```

Set the seed for reproducibility:

```{r set-seed}
set.seed(1)
```

Simulate a set of noisy ARIMA signals:

```{r sim-amira}
n <- 1024
m <- 100
SNR <- 2
v <- DJ.EX()
x <- seq(1,n)/n
ssig <- sd(v$bumps)
sigma <- ssig/SNR
e <- arima.sim(n = n,model = list(ar = 0.99,ma = 1))
Y <- matrix(0,m,n)
for (i in 1:m) {
  e <- sigma*e/sqrt(var(e))
  y <- v$bumps + e
  Y[i,] <- y
}
```

This is the first noisy signal:

```{r plot-amira, fig.height=2, fig.width=3.5}
y <- Y[1,]
pdat <- data.frame(x = x,y = y)
ggplot(pdat,aes(x = x,y = y)) +
  geom_line() +
  theme_cowplot(font_size = 12)
```

Now let's compute the discrete wavelet transform (DWT) for each
signal:

```{r wd}
D <- matrix(0,m,n-1)
for (i in 1:m) {
  y <- Y[i,]
  D[i,] <- wd(y)$D
}
```

One way to illustrate the decorrelating effect of the wavelet
transform is to compare the autocorrelation for the original signal
and for the wavelet coefficients (WCs). Let's try this on the first
noisy signal:

```{r acf, fig.height=3, fig.width=6}
par(mfrow = c(1,2))
y <- Y[1,]
d <- D[1,]
acf(y)
acf(d)
```

Observe that the autocorrelation is mostly near zero for the WCs
(right).


