---
title: Illustrate whitening property of wavelets
author: Peter Carbonetto
output: workflowr::wflow_html
---

Here we give two simple examples to illustrate the decorrelating
("biwhitening") effect of the wavelet transform. These examples use
synthetic signals not generated using wavelets.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load the wavethresh package as well as a few other packages used this
demo:

```{r load-pkgs, message=FALSE, results="hide"}
library(wavethresh)
library(reshape)
library(ggplot2)
library(cowplot)
```

Set the seed for reproducibility:

```{r set-seed}
set.seed(1)
```

1. Bumps example
----------------

Simulate a set of signals using the "bumps" test function from Donoho
& Johnson, with ARMA noise:

```{r sim-amira}
n <- 1024
m <- 100
SNR <- 2
v <- DJ.EX()
x <- seq(1,n)/n
ssig <- sd(v$bumps)
sigma <- ssig/SNR
Y <- matrix(0,m,n)
for (i in 1:m) {
  e <- arima.sim(n = n,model = list(ar = 0.99,ma = 1))
  e <- sigma*e/sqrt(var(e))
  y <- v$bumps + e
  Y[i,] <- y
}
```

This is was the first noisy signal looks like:

```{r plot-amira, fig.height=2, fig.width=3.5}
y <- Y[1,]
pdat <- data.frame(x = x,y = y)
ggplot(pdat,aes(x = x,y = y)) +
  geom_line() +
  theme_cowplot(font_size = 12)
```

Now let's compute the discrete wavelet transform (DWT) for each
signal:

```{r wd}
D <- matrix(0,m,n-1)
for (i in 1:m) {
  y <- Y[i,]
  D[i,] <- wd(y)$D
}
```

One way to illustrate the decorrelating effect of the wavelet
transform is to compare the autocorrelation for the original signal
and for the wavelet coefficients (WCs). Let's try this on the first
noisy signal:

```{r acf, fig.height=3, fig.width=6}
par(mfrow = c(1,2))
y <- Y[1,]
d <- D[1,]
acf(y)
acf(d)
```

Observe that the autocorrelation is mostly near zero for the WCs
(right-hand plot).

Another way to illustrate the decorrelating effect of the DWT is to
compute the sample correlations. Compare the sample correlations in
the original data space to the sample correlations in the wavelet
space:

```{r cor, fig.height=4, fig.width=3, warning=FALSE}
cY <- cor(Y)
cD <- cor(D)
p1 <- ggplot(data.frame(x = cY[upper.tri(cY)]),aes(x = x)) +
  geom_histogram(fill = "black",color = "white",bins = 64) +
  xlim(-1,1) +
  labs(x = "correlation",title = "data space") +
  theme_cowplot(font_size = 12)
p2 <- ggplot(data.frame(x = cD[upper.tri(cD)]),aes(x = x)) +
  geom_histogram(fill = "black",color = "white",bins = 64) +
  xlim(-1,1) +
  labs(x = "correlation",title = "wavelet space") +
  theme_cowplot(font_size = 12)
plot_grid(p1,p2,nrow = 2,ncol = 1)
```

Indeed, the strong correlations in the original data space are removed
in the wavelet space.

2. Gaussian process example
---------------------------

In the second example, we simulate a small data set from a Gaussian
process (GP) with a Ornstein-Uhlenbeck (O-H) kernel:

```{r sim-gp}
n <- 2000
x <- seq(1,32)
d <- as.matrix(dist(x))
Sigma <- exp(-d/30)
Y <- mvrnorm(n,rep(0,32),Sigma)
```

In a GP model, the correlations between the points "decay" with
distance:

```{r plot-cor-gp, fig.height=3.25, fig.width=4, warning=FALSE}
pdat <- melt(cor(Y))
colnames(pdat) <- c("x","y","cor")
ggplot(pdat,aes(x = x,y = y,fill = cor)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  theme_cowplot(font_size = 12)
```

Now perform the discrete wavelet transform on these data:

```{r wd-gp}
D <- t(apply(Y,1,function (x) wd(x)$D))
```

Examining the first signal, the autocorrelation is very strong in the
original data space, and almost completely vanishes after applying the
DWT:

```{r acf-gp, fig.height=3, fig.width=6}
par(mfrow = c(1,2))
y <- Y[1,]
d <- D[1,]
acf(y)
acf(d)
```

Likewise, the original correlations are very strong, whereas the
correlations among the WCs are mostly zero or close to zero:

```{r cor-gp, fig.height=4, fig.width=3, warning=FALSE}
cY <- cor(Y)
cD <- cor(D)
p1 <- ggplot(data.frame(x = cY[upper.tri(cY)]),aes(x = x)) +
  geom_histogram(fill = "black",color = "white",bins = 64) +
  xlim(-1,1) +
  labs(x = "correlation",title = "data space") +
  theme_cowplot(font_size = 12)
p2 <- ggplot(data.frame(x = cD[upper.tri(cD)]),aes(x = x)) +
  geom_histogram(fill = "black",color = "white",bins = 64) +
  xlim(-1,1) +
  labs(x = "correlation",title = "wavelet space") +
  theme_cowplot(font_size = 12)
plot_grid(p1,p2,nrow = 2,ncol = 1)
```
